---
title: "RTqPCR_CryptoMobius"
author: "Liz Hughes"
date: "12/01/23"
output:
  html_document:
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE , warning=FALSE, message=FALSE)
```

# Experimental Design

This experiment was set up to examine the expression of Gat201 and some Gat201 target genes in different strains grown under the same conditions.

RTqPCR of WT, Gat201 deletion, Complemented clones C26 and C30. 

This experiment was set up to show the complementation of native Gat201 into SH4 in the Madahani GAT201 deletion strain (C26 and C30).

Strains: 
H99, KN99-alpha, Madhani Gat201 deletion mutant, and 2 Gat201 complemented strains (into Madhani-Gat201 deletion strain) C26 and C30.

Culture conditions:
Set up overnight cultures from a single colony in YPD at 30 degrees, 150rpm, 3 biological replicates.
Cultures were set up in 20 ml RPMI at an OD of 0.1 from the ON culture. Cell were washed with RPMI prior to testing the OD and setting up cultures in RPMI. Cultures were incubated at 37 degrees, 150rpm for 7 hours.

Sample collection: 
19 ml Samples were collected at 7 hours. Samples were fixed in methanol on dry ice for 3-5 mins. My Cryptococcus RNA extraction protocol was followed to isolate RNA from samples.

Replicates:
3 biological replicates were collected for each sample.
3 technical replicates +RT and 1 -RT

Test Primers:
Plate 1:
Gat201 CNAG_01551 (set 1)
PDR802 CNAG_03894 (Set 5)
TSA3 CNAG_06917 (Set 2)

Plate 2: 
BLP1  CNAG_06346 (Set 4)
Meu1  CNAG_00165 (Set 4)
MEP1  CNAG_04735 (Set 1)


Reference gene primers for both plates: 
SRP14 CNAG_04994 (Set 5)
ACT1  CNAG_00483 (Set 5) 
GPD  CNAG_00121 (Set 5)

# Load Packages

```{r load_packages}

library(tidyverse)
library(tidyr)
library(cowplot)
library(tidyqpcr)
library(dplyr)
library(rlang)
library(markdown)
library(ggplot2)
library(here)

# set default theme for plots

theme_set(theme_bw(base_size = 11) %+replace%
  theme(
    strip.background = element_blank(),
    panel.grid = element_blank()
  ))
```

# Plate 1
## Create Plateplan 1
### List target_ids of primer sets

```{r target_ids_plate1}

colkey <- create_colkey_6_in_24(target_id = c("Gat201", "TSA3", "PDR802", "SRP14", "ACT1", "GPD"))

print(colkey)
```

## Set up experimental samples plate 1
### List sample name, BioRep (3) and TechRep (3)

```{r samples_plate1}

Samples <- c("H99-1", "H99-2", "H99-3", "KN99alpha-1", "KN99alpha-2", "KN99alpha-3", "deltaGat201-1", "deltaGat201-2", "deltaGat201-3", "C26-1", "C26-2", "C26-3", "C30-1", "C30-2", "C30-3")

TechReps <- c("1", "2", "3")

rowkey <- tibble(well_row=LETTERS[1:15], 
                 sample_id= Samples)

print(rowkey)
```

## Plateplan1

```{r plate_plan1}

plate1 <- label_plate_rowcol(
  plate = create_blank_plate(well_row = LETTERS[1:15], well_col = 1:24),
  rowkey = rowkey,
  colkey = colkey
)
print(plate1)
display_plate_qpcr(plate1)
```

## Load data for plate 1

```{r load_data_plate1}

plate1 <-  read_tsv("../data/EH_20230112_CryptoMob_plate1_Ct.txt", 
            skip=1) %>%
          mutate(well=Pos,cq=Cp) %>%
          right_join(plate1)

plate1
```

## Split Sample column into BioRep, Time, Condition plate 1

```{r separate_conditions_plate1}

plate1_sep <- separate(plate1, 
         sample_id, 
         into = c("Sample" , "BioRep"), 
         sep = "-", 
         remove = FALSE)

```

## Plot all reps, unnormalized plate 1

```{r plot_plate1_sep}

ggplot(data = plate1_sep) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = Sample),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized"
  ) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Check reference genes plate 1

```{r plate_sep_refgenes_plate1}

  ggplot(data = filter(plate1_sep, target_id == c("SRP14", "ACT1", "GPD") )) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized"
  ) +
  facet_grid(~Sample) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

ACT1 has a lower Ct value then SRP14 and GPD, indicating higher mRNA expression in these conditions.

## Normalize data with 3 reference genes plate 1 
### Within sample normalizing

```{r plate1_norm}

plate1_norm <- plate1_sep %>%
  group_by(sample_id)%>%
  filter(prep_type == "+RT" )%>%
  calculate_deltacq_bysampleid(ref_target_ids = c("SRP14", "ACT1", "GPD"))

plate1_norm
```

## Normalise data with ACT1 reference gene only plate 1

Test normalization with single reference gene since expression of AC T1 is greater than the other 2 reference genes.

```{r plate1_norm_ACT1}

plate1_norm_ACT1 <- plate1_sep %>%
  group_by(sample_id)%>%
  filter(prep_type == "+RT" )%>%
  calculate_deltacq_bysampleid(ref_target_ids = c( "ACT1"))

plate1_norm_ACT1

```


## Plot delta Cq for normalized data plate1 ref genes

Use these data frames to plot with and without reference genes:
"plate1_norm" for all genes and strains
"plot1" for non-ref genes only without H99
"plotr" for ref genes only


```{r select_samples_plot1}

plot1 <- (data = filter(plate1_norm, target_id != c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))
```

```{r select_samples_plot1r}
plot1r <- (data = filter(plate1_norm, target_id == c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))

```

## Plot delta Cq for normalized data 

```{r Plate1_norm_plot_Cq}

ggplot(data = plot1r ) +
  geom_point(aes(x = target_id, y = delta_cq, shape = Sample, colour = BioRep ),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  
  labs(x = "Target Gene", 
       y = "Normalised Delta Cq",
       title = "Expression of Ref Genes only") +
  facet_wrap(~Sample, ncol = 5)+
    theme(text = element_text(size = 10)) +
    theme(axis.text.x = element_text(size = 8, angle = 90))+
    theme(axis.text.y = element_text(size= 10))+
    scale_color_brewer(palette="Dark2")


```
"plate1_norm": GPD and SRP14 have a Cq=0 in all samples. ACT1 has a Cq = -5 in all samples.

"plot1": deltaGat201 has a higher Cq value, indicating less starting material. In all strains PDR802 and TSA3 show similar Cq values.

"plot1r": Looking at the refernce genes alone you can see that ACT1 is expressed much higher that SRP14 and GPD. 

Maybe should try normalising with SRP14 and GPD only?


## Save plots with all genes, only target genes and only reference genes plate 1

Change "ggplot(data = plot1)" to make each plot and change in plot name below before saving.
"plate1_norm" for all genes and strains
"plot1" for non-ref genes only without H99
"plot1r" for ref genes only

```{r save_plot1_jpeg}

ggsave("../Results/plot1r.png", 
 ,width = 10,height=7)
```
`
## Calculate a summary value for each sample-target combination

```{r plate1_med}

plate1_med <- plate1_norm %>%
  group_by(sample_id, Sample, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE)
  )

plate1_med
```

### Filter plate1_med for individual genes Gat201, PDR802 and TSA3

Filtering by gene makes it easier to digest the data.

```{r plate1_med_filter}

filter(plate1_med, target_id == "Gat201")
filter(plate1_med, target_id == "PDR802")
filter(plate1_med, target_id == "TSA3")
```
This confirms that deltaGat201 has far less rel_abundance of each gene compared to the other strains.
PDR802 has a similar rel_abundance for all  strains.
Both WT strains H99 and KN99-alpha have the highest rel_abundance of all 3 genes, followed by C30and C26, then deltaGat201.

## Plot relative abundance for normalized data plate 1 

```{r plot1_relabun}

ggplot(data = filter(plate1_med, target_id != c("SRP14", "ACT1", "GPD"))) + 
  geom_point(aes(x = target_id, y = rel_abund, shape = Sample, colour = sample_id)) + 
  labs(   y = "RNA abundance relative to Ref Genes" ) +
  facet_wrap(~Sample, ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Plot delta Cq for normalized data with ACT1 only plate 1

```{r select_samples_plate1_ACT1}

plot1_ACT1 <- (data = filter(plate1_norm_ACT1, target_id != c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))
```


```{r Plate1_norm_ACT1_plot_Cq}

ggplot(data = plot1_ACT1) +
  geom_point(aes(x = target_id, y = delta_cq, shape = Sample, colour = BioRep ),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  
  labs(x = "Target Gene", 
       y = "Normalised Delta Cq",
       title = "Expression of GAT201 and Targets_ACT1 only") +
    theme(text = element_text(size = 12)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))+
    scale_color_brewer(palette="Dark2")


```


## Save plot with only target genes normalised by ACT1 only

```{r save_plot1_ACT1_jpeg}

ggsave("../Results/Plate1_norm_ACT1.png", 
 ,width = 10,height=7)
```

## Doubly normalized data: delta delta Cq plate 1
### Calculate delta delta cq against a chosen reference sample KN99-alpha (WT)

This is an estimate of log2 fold change. 
Compare WT to other strains in the same growth conditions

```{r plate1_deltanorm}

plate1_deltanorm <- plate1_med %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = "KN99alpha-1")

plate1_deltanorm_med <- plate1_deltanorm %>%
    group_by(sample_id, Sample, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE)
  )

plate1_deltanorm
plate1_deltanorm_med
```

## Plot delta delta Cq (log2-fold change) for each target gene plate 1

Here, delta delta Cq is positive when a target is more highly detected in the relevant sample, compared to reference samples.

```{r ddCq1}

ggplot(data = filter(plate1_deltanorm_med, target_id != "SRP14") ) + 
  geom_line(aes(x = target_id, y = deltadelta_cq, colour = sample_id, group = sample_id)) + 
  geom_point(aes(x = target_id, y = deltadelta_cq,  colour = sample_id)) + 
  labs(
    y = "delta delta Cq (log2 fold change)\n relative to H99"
  ) +
    facet_wrap(~Sample, ncol=5) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8))+
    theme(text = element_text(size = 9)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))

```

## Save delta delta Cq (log2 fold change)\n relative to H99 plate 1

```{r save_ddcq_plot1_jpeg}

ggsave("../Results/ddcq_plate1.png", 
 width = 10,height=7)

```

## Plot fold change for each target gene plate 1

```{r foldchange_plate1}

ggplot(data = filter(plate1_deltanorm_med, target_id != "SRP14") ) + 
  geom_line(aes(x = target_id, y = fold_change, colour = sample_id, group = sample_id)) + 
  geom_point(aes(x = target_id, y = fold_change, colour = sample_id)) + 
  labs(
    y = "fold change relative to KN99alpha-1"
  ) +
  facet_wrap(~Sample,ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8))+
    theme(text = element_text(size = 9)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))

```

## Save fold change relative to KN99alpha-1 plot1

```{r save_fc_plot1_jpeg}

ggsave("../Results/fc_plate1.png", 
 width = 10,height=7)
```


# Plate 2
## Create Plateplan 2
### list target_ids of primer sets

```{r target_ids_plate2}

colkey <- create_colkey_6_in_24(target_id = c("BLP1", "MEU1", "MEP1", "SRP14", "ACT1", "GPD"))

print(colkey)
```

## Set up experimental samples plate 2
### List sample name, BioRep (3) and TechRep (3)

```{r samples_plate2}

Samples <- c("H99-1", "H99-2", "H99-3", "KN99alpha-1", "KN99alpha-2", "KN99alpha-3", "deltaGat201-1", "deltaGat201-2", "deltaGat201-3", "C26-1", "C26-2", "C26-3", "C30-1", "C30-2", "C30-3")

TechReps <- c("1", "2", "3")

rowkey <- tibble(well_row=LETTERS[1:15], 
                 sample_id= Samples)

print(rowkey)
```

## Plateplan2

```{r plate_plan2}

plate2 <- label_plate_rowcol(
  plate = create_blank_plate(well_row = LETTERS[1:15], well_col = 1:24),
  rowkey = rowkey,
  colkey = colkey
)

print(plate2)
display_plate_qpcr(plate2)

```

## Load data for plate 2

```{r load_data_plate2}

plate2 <- read_tsv( "../data/EH_20230112_CryptoMob_plate2_Ct.txt",  
                    skip=1) %>%
    mutate(well=Pos,cq=Cp) %>%
    right_join(plate2)

```


## Split Sample column into BioRep, Time, Condition plate 2

```{r separate_conditions_plate2}

plate2_sep <- separate(plate2, 
         sample_id, 
         into = c("Sample" , "BioRep"), 
         sep = "-", 
         remove = FALSE)

```

## Plot all reps, unnormalized plate 2

```{r plot_plate2_sep}

ggplot(data = plate2_sep) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized"
  ) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Check reference genes plate 2

```{r plate_sep_refgenes_plate2}

  ggplot(data = filter(plate2_sep, target_id == c("SRP14", "ACT1", "GPD") )) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized"
  ) +
  facet_grid(~sample_id) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Normalize data with 3 reference genes plate 2
### Within sample normalizing

```{r plate2_norm}

plate2_norm <- plate2_sep %>%
  group_by(sample_id)%>%
  filter(prep_type == "+RT" )%>%
  calculate_deltacq_bysampleid(ref_target_ids = c("SRP14", "ACT1", "GPD"))

```

## Normalise data with ACT1 reference gene only plate 2

```{r plate2_norm_ACT1}

plate2_norm_ACT1 <- plate2_sep %>%
  group_by(sample_id)%>%
  filter(prep_type == "+RT" )%>%
  calculate_deltacq_bysampleid(ref_target_ids = c( "ACT1"))

```


## Plot delta Cq for normalized data plate2 ref genes
### plate2_norm for all genes and strains
### plot2 for non-ref genes only without H99
### plot2r for ref genes only

```{r select_samples_plot2}

plot2 <- (data = filter(plate2_norm, target_id != c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))
```


```{r select_samples_plot1plot2r}
plot2r <- (data = filter(plate2_norm, target_id == c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))

```


```{r Plate2_norm_plot_Cq}

ggplot(data = plate2_norm ) +
  geom_point(aes(x = target_id, y = delta_cq, shape = Sample, colour = BioRep ),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  
  labs(x = "Target Gene", 
       y = "Normalised Delta Cq",
       title = "Expression of GAT201 and Targets") +
  facet_wrap(~Sample, ncol = 5)+
    theme(text = element_text(size = 11)) +
    theme(axis.text.x = element_text(size = 8, angle = 90))+
    theme(axis.text.y = element_text(size= 10))+
    scale_color_brewer(palette="Dark2")

```

The reference genes show the same expression as plate 1 with ACT1 being expressed higher than SRP14 and GPD.

The target gene BLP1 is expressed similarly in deltaGat201 and the complemented strains. It is expressed More in KN99alpha.
This is the same for MEP1 and MEU1. This suggests the complementation is not acting on GAT201 target genes in the same way as wild type. This could be because expression of GAT201 is less in the complemented strains compared to WT.

## Save plots with all genes, only target genes and only reference genes plate 2

```{r save_plot2_jpeg}

ggsave("../Results/plate2_norm.png", 
 ,width = 10,height=7)
```

`
## Calculate a summary value for each sample-target combination in plate 2

```{r plate2_med}

plate2_med <- plate2_norm %>%
  group_by(sample_id, Sample, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE)
  )

```

### Filter plate2_med for individual genes

```{r plate2_med_filter}

filter(plate2_med, target_id == "BLP1")
filter(plate2_med, target_id == "MEU1")
filter(plate2_med, target_id == "MEP1")
```

BLP1, MEU1 and MEp 1 have a rel abundance greater in wild type cells compared to deltaGat201 and the complemented strains.

## Plot relative abundance for normalized data plate 2 ref genes

```{r plot2_relabun}

ggplot(data = filter(plate2_med, target_id != c("SRP14", "ACT1", "GPD"))) + 
  geom_point(aes(x = target_id, y = rel_abund, shape = Sample, colour = sample_id)) + 
  labs(   y = "RNA abundance relative to Ref Genes" ) +
  facet_wrap(~Sample, ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```


## Plot delta Cq for normalized data with ACT1 only plate 2

```{r select_samples_plate2_ACT1}

plot2_ACT1 <- (data = filter(plate2_norm_ACT1, target_id != c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))
```


```{r Plate2_norm_ACT1_plot_Cq}

ggplot(data = plot2_ACT1) +
  geom_point(aes(x = target_id, y = delta_cq, shape = Sample, colour = BioRep ),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  
  labs(x = "Target Gene", 
       y = "Normalised Delta Cq",
       title = "Expression of GAT201 and Targets") +
    theme(text = element_text(size = 12)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))+
    scale_color_brewer(palette="Dark2")


```


## Save plot normalized data with ACT1 only plate 2

```{r save_plot2_ACT1_jpeg}

ggsave("../Results/plate2_norm_ACT1.png", 
 ,width = 10,height=7)
```

## Doubly normalized data: delta delta Cq plate 2
### Calculate delta delta cq against a chosen reference sample KN99-alpha (WT)

This is an estimate of log2 fold change. 
plate2
Compare WT to other strains in the same growth conditions

```{r plate2_deltanorm}

plate2_deltanorm <- plate2_med %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = "KN99alpha-1")

plate2_deltanorm_med <- plate2_deltanorm %>%
    group_by(sample_id, Sample, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE)
  )

plate2_deltanorm
plate2_deltanorm_med
```

## Plot delta delta Cq (log2-fold change) for each target gene in plate 2

Here, delta delta Cq is positive when a target is more highly detected in the relevant sample, compared to reference samples.

```{r ddCq2}

ggplot(data = filter(plate2_deltanorm_med, target_id != "SRP14") ) + 
  geom_line(aes(x = target_id, y = deltadelta_cq, colour = sample_id, group = sample_id)) + 
  geom_point(aes(x = target_id, y = deltadelta_cq,  colour = sample_id)) + 
  labs(
    y = "delta delta Cq (log2 fold change)\n relative to H99"
  ) +
  facet_wrap(~Sample, ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Save delta delta Cq (log2 fold change)\n relative to H99 plate 2

```{r save_ddcq_plot2_jpeg}

ggsave("../Results/ddcq_plate2.png", 
 width = 10,height=7)

```


## Plot fold change for each target gene plate 2

```{r foldchange_plate2}

ggplot(data = filter(plate2_deltanorm_med, target_id != "SRP14") ) + 
  geom_line(aes(x = target_id, y = fold_change, colour = sample_id, group = sample_id)) + 
  geom_point(aes(x = target_id, y = fold_change, colour = sample_id)) + 
  labs(
    y = "fold change relative to KN99alpha-1"
  ) +
  facet_wrap(~Sample,ncol=5) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Save fold change relative to KN99alpha-1 plot2

```{r save_fc_plot2_jpeg}

ggsave("../Results/fc_plate2.png", 
 width = 10,height=7)
```


# Combine plates together

```{r bind_plates}
plates <- bind_rows(plate1_norm, plate2_norm) 

```

## Plot delta Cq for normalized data combined plates
### plates for all genes and strains (normalised)
### plot_comb for non-ref genes only without H99
### plot_combr for ref genes only

```{r select_samples_comb}

plot_comb <- (data = filter(plates, target_id != c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))
```


```{r select_samples_combr}
plot_combr <- (data = filter(plates, target_id == c("SRP14", "ACT1", "GPD"), Sample != c("H99" )))

```


```{r Plates_plot_Cq}

ggplot(data = plates) +
  geom_point(aes(x = target_id, y = delta_cq, shape = Sample, colour = BioRep ),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  
  labs(x = "Target Gene", 
       y = "Normalised Delta Cq",
       title = "Expression of GAT201 and Targets") +
    theme(text = element_text(size = 12)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))+
    scale_color_brewer(palette="Dark2")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))


```

## Save combined plots with all genes, target genes only and reference genes only

```{r save_comb_plot_jpeg}

ggsave("../Results/plates.png", 
 ,width = 10,height=7)
```

## Calculate a summary value for each sample-target combination in combined plates

```{r plates_med}

plates_med <- plates %>%
  group_by(sample_id, Sample, target_id) %>%
  summarize(
    delta_cq  = median(delta_cq, na.rm = TRUE),
    rel_abund = median(rel_abund, na.rm = TRUE)
  )

```

### Filter plates_med for individual genes

```{r plates_med_filter}

filter(plates_med, target_id == "Gat201")
filter(plates_med, target_id == "PDR802")
filter(plates_med, target_id == "TSA3")
filter(plates_med, target_id == "BLP1")
filter(plates_med, target_id == "MEU1")
filter(plates_med, target_id == "MEP1")
```

## Plot relative abundance for normalized data combined plates ref genes

```{r plates_relabun}

ggplot(data = filter(plates_med, target_id != c("SRP14", "ACT1", "GPD"))) + 
  geom_point(aes(x = target_id, y = rel_abund, shape = Sample, colour = target_id)) + 
  labs(   y = "RNA abundance relative to Ref Genes" ) +
  facet_wrap(~Sample, ncol=5) + 
 theme(text = element_text(size = 10)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))+
    scale_color_brewer(palette="Dark2")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size =7))

  

```

## Doubly normalized data: delta delta Cq combined plates
### Calculate delta delta cq against a chosen reference sample KN99-alpha (WT)

This is an estimate of log2 fold change. 
Combined plates
Compare WT to other strains in the same growth conditions

```{r plates_deltanorm}

plates_deltanorm <- plates_med %>%
  calculate_deltadeltacq_bytargetid(ref_sample_ids = "KN99alpha-1")

plates_deltanorm_med <- plates_deltanorm %>%
    group_by(sample_id, Sample, target_id) %>%
  summarize(
    deltadelta_cq  = median(deltadelta_cq, na.rm = TRUE),
    fold_change    = median(fold_change,   na.rm = TRUE)
  )

plates_deltanorm
plates_deltanorm_med
```

## Plot delta delta Cq (log2-fold change) for each target gene in combined plates

Here, delta delta Cq is positive when a target is more highly detected in the relevant sample, compared to reference samples.

```{r ddCq_plates}

ggplot(data = filter(plates_deltanorm_med, target_id != "SRP14") ) + 
  geom_line(aes(x = target_id, y = deltadelta_cq, colour = sample_id, group = sample_id)) + 
  geom_point(aes(x = target_id, y = deltadelta_cq,  colour = sample_id)) + 
  labs(
    y = "delta delta Cq (log2 fold change)\n relative to H99"
  ) +
  facet_wrap(~Sample, ncol=5) + 
 theme(text = element_text(size = 10)) +
    theme(axis.text.x = element_text(size = 10))+
    theme(axis.text.y = element_text(size= 10))+
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size= 7))


```


## Save delta delta Cq (log2 fold change)\n relative to H99 combined plates

```{r save_ddcq_plots_jpeg}

ggsave("../Results/ddcq_plates.png", 
 width = 10,height=7)

```

## Plot fold change for each target gene combined plates

```{r foldchange_plates}

ggplot(data = filter(plates_deltanorm_med, target_id != "SRP14") ) + 
  geom_line(aes(x = target_id, y = fold_change, colour = sample_id, group = sample_id)) + 
  geom_point(aes(x = target_id, y = fold_change, colour = sample_id)) + 
  labs(
    y = "fold change relative to KN99alpha-1"
  ) +
  facet_wrap(~Sample,ncol=5) + 
 theme(text = element_text(size = 10)) +
    theme(axis.text.x = element_text(size = 11))+
    theme(axis.text.y = element_text(size= 11))+
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size= 7))

```

## Save fold change combined plates

```{r save_fc_plots_jpeg}

ggsave("../Results/fc_plates.png", 
 width = 10,height=7)

```


