---
title: "RTqPCR_CryptoMobius"
author: "Liz Hughes"
date: "16/12/2022"
output:
  html_document:
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , warning=FALSE, message=FALSE)
```

# Experimental Design

This experiment was set up to examine the expression of Gat201 in different strains grown under the same conditions.

RTqPCR of WT, Gat201 deletion, Complemented clones C26 and C30. 

This experiment was set up to show the complementation of native Gat201 into SH4 in the Madahani GAT201 deletion strain.

Strains: 
KN99-alpha (KNa), Madhani-Gat201 deletion mutant (M-Gat201) and 2 Gat201 complemented strains (into Madhani-Gat201 deletion strain) C26 and C30.

Culture conditions:
Set up overnight cultures from a single colony in YPD at 30 degrees, 150rpm, 3 biological replicates.
Cultures were set up in 20 ml RPMI at an OD of 0.1 from the ON culture. Cell were washed with RPMI prior to testing the OD and setting up cultures in RPMI. Cultures were incubated at 37 degrees, 150rpm for 5 hours.

Sample collection: 
19 ml Samples were collected at 5 hours. Samples were fixed in methanol on dry ice for 3-5 mins. My Cryptococcus RNA extraction protocol was followed to isolate RNA from samples.

Replicates:
3 biological replicates were collected for each sample.
3 technical replicates +RT and 1 -RT

Test Primers:
Plate 1:
Gat201 CNAG_01551 (set 1)
PDR802 CNAG_03894 (Set 1)
TSA3 CNAG_06917 (Set 1)

Plate 2: 
BLP1  CNAG_06346
Meu1  CNAG_00165
MEP1  CNAG_04735


Reference gene primers for both plates: 
SRP14  CNAG_04994 
ACT1  CNAG_00483 (Set 1) 
GDP1  CNAG_00121 (Set 1)



```{r load_packages}

library(tidyverse)
library(cowplot)
library(tidyqpcr)
library(dplyr)
library(rlang)
library(markdown)
library(ggplot2)
theme_set(theme_bw())
```

# Create Plateplan 1
## list target_ids of primer sets

```{r target_ids_plate1}

colkey <- create_colkey_6_in_24(target_id = c("Gat201", "TSA3", "PDR802", "SRP14", "ACT1", "GDP1"))

print(colkey)
```

# Set up experimental samples
##List sample name, BioRep (3) and TechRep (3)
```{r samples}
Samples <- c("H99-1", "H99-2", "H99-3", "KN99alpha-1", "KN99alpha-2", "KN99alpha-3", "deltaGat201-1", "deltaGat201-2", "deltaGat201-3", "C26-1", "C26-2", "C26-3", "C30-1", "C30-2", "C30-3")

TechReps <- c("1", "2", "3")

rowkey <- tibble(well_row=LETTERS[1:15], 
                 sample_id= Samples)

print(rowkey)
```


## Plateplan1
```{r plate_plan1}

plate1 <- label_plate_rowcol(
  plate = create_blank_plate(well_row = LETTERS[1:15], well_col = 1:24),
  rowkey = rowkey,
  colkey = colkey
)

print(plate1)
display_plate(plate1)

```




# Load data for plate 1

```{r load_data_plate1}

plate1 <- read_tsv( "../Input/20221014_ComplementCheck_EH_Ct.txt",  
                    skip=1) %>%
    mutate(well=Pos,cq=Cp) %>%
    right_join(plate1)

plate1
```


# Split Sample column into BioRep, Time, Condition

```{r separate_conditions_plate1}

plate1_sep <- separate(plate1, 
         sample_id, 
         into = c("Sample" , "BioRep"), 
         sep = "-", 
         remove = FALSE)
view(plate1_sep)
```


# Plot all reps, unnormalized


```{r plot_plate1_sep}

ggplot(data = plate1_sep) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized"
  ) +
  facet_wrap(~sample_id) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

# Check reference genes plate 1


```{r plate_sep_refgenes_plate1}

  ggplot(data = filter(plate1_sep, target_id == c("SRP14", "ACT1", "GDP1") )) +
  geom_point(aes(x = target_id, y = cq, shape = prep_type, colour = target_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(
    y = "Quantification cycle (Cq)",
    title = "All reps, unnormalized"
  ) +
  facet_grid(~sample_id) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

# Normalise data with 3 reference genes


```{r plate1_norm}

plate1_norm <- plate1_sep %>%
  group_by(sample_id)%>%
  filter(prep_type == "+RT" )%>%
  calculate_deltacq_bysampleid(ref_target_ids = c("SRP14", "ACT1", "GDP1"))

plate1_norm

```

# Plot delta Cq for normalized data

```{r select_samples}

plot1 <- (data = filter(plate1_norm, target_id != c("SRP14", "ACT1", "GDP1"), Sample != c("H99" )))
```



```{r Plate1_norm_plot_Cq}

ggplot(data = plot1) +
  geom_point(aes(x = target_id, y = delta_cq, shape = Sample, colour = BioRep ),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  
  labs(x = "Target Gene", 
       y = "Normalised Delta Cq",
       title = "Expression of GAT201 and Targets") +
    theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_text(size = 14))+
    theme(axis.text.y = element_text(size= 14))+
    scale_color_brewer(palette="Dark2")


```


```{r save_plot1_jpeg}

ggsave("../Results/Clone_Plate1.png", 
 ,width = 10,height=7)
```


# Plot data only for Gat201

```{r Plate_norm_plot_Cq_Gat201}

ggplot(data = filter(plate_norm, target_id == c("Gat201") )) +
  geom_point(aes(x = target_id, y = delta_cq, colour = sample_id),
    position = position_jitter(width = 0.2, height = 0)
  ) +
  labs(y = "delta Cq") +
  facet_grid(~sample_id) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))


```


